---
title: "US Births 2000-2014"
author: "GS"
date: "18/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, comment = "", message = FALSE, fig.align = "center")
```

```{r}
# USbirths2000-2014_SSA.csv` contains U.S. births data for the years 2000 to 2014, as provided by the Social Security Administration.
# 
# Header                    Definition
# ------                    ----------
# 
# year                      Year
# month	                    Month
# date_of_month	            Day number of the month
# day_of_week	Day of week   where 1 is Monday and 7 is Sunday
# births	                  Number of births
```

[Link Kaggle](https://www.kaggle.com/mysarahmadbhat/us-births-2000-to-2014?select=US_births_2000-2014_SSA.csv)

```{r, packages}
library(tidyverse)

theme_set(theme_bw())
```

```{r, dataset}
births <- read_csv("US_births_2000-2014_SSA.csv")

sample_n(births, 5)
```


```{r, cleaning}
births <- births %>% 
  mutate(year = as_factor(year)) %>% 
  mutate(month = as_factor(month)) %>% 
  mutate(date_of_month = as_factor(date_of_month)) %>% 
  mutate(day_of_week = as_factor(day_of_week))  

levels(births$month) <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")
levels(births$day_of_week) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

births
```
```{r, births}
ggplot(births, aes(births)) + 
  geom_histogram(bins = sqrt(nrow(births)), color = "midnightblue", fill = "white") +
  labs(title = "Distribution of births in US",
       subtitle = "2000 - 2014",
       x = "Number of births",
       y = "Count of days")

ggplot(births, aes(births, color = year)) + 
  geom_density() +
  labs(title = "Distribution of births in US",
       subtitle = "2000 - 2014",
       x = "Number of births",
       y = "Count of days")

summary(births$births)
```

It seems to be two means in our dataset (binomial)

```{r}
ggplot(births, aes(y=births, x = year, color = year)) + 
  geom_boxplot() +
  labs(title = "Distribution of births in US",
       subtitle = "2000 - 2014",
       x = "Number of births",
       y = "Count of days") + 
  theme(legend.position = "none")

```
It seems to be an insignificant increase of births in 2006 and 2007 

```{r, month}
ggplot(births, aes(births, color = year)) + 
  geom_density() +
  facet_wrap(~month)
  labs(title = "Distribution of births in US",
       subtitle = "2000 - 2014",
       x = "Number of births",
       y = "Count of days")
```
After splitting the dataset by year and month I see the same binomial behavior

```{r, day_of_week}
ggplot(births, aes(births, color = year)) + 
  geom_density() +
  facet_wrap(~day_of_week)
  labs(title = "Distribution of births in US",
       subtitle = "2000 - 2014",
       x = "Number of births",
       y = "Count of days")
```

After splitting the dataset by day_of_week it looks like on Saturdays and Sundays are less children born

```{r, full, cache=TRUE}
ggplot(births, aes(births, color = year)) + 
  geom_density() +
  facet_grid(day_of_week~month)
  labs(title = "Distribution of births in US",
       subtitle = "2000 - 2014",
       x = "Number of births",
       y = "Count of days")
```
I see the same behavior if I split the dataset by year, day_of_week and month.  There are some peaks around 12500 on Fridays

I want see if this difference between weekends and weekdays is significant.  Even less children born on Sundays than on Saturdays.  
There are almost 50% of more births during weekdays than during weekends with a significant p-value.
I thought I could find a difference among weekdays (most on Thursdays and Fridays) but it was not the case.
```{r, weekend weekdays}
births <- births %>% 
  mutate(weekend = if_else(day_of_week == "Sunday" | day_of_week == "Saturday", "Yes", "No"))


births %>% 
  group_by(weekend) %>% 
  summarise(mean = mean(births))

births %>% 
  group_by(day_of_week) %>% 
  summarise(mean = mean(births)) 


library(infer)

t_test(x = births, 
       formula = births ~ weekend, 
       order = c("Yes", "No"),
       alternative = "two-sided")
```


## Conclusion
Weekends born less children than in weekdays (p-value is significant).  Even lees children born on Sundays.
Potentially, the physicians want to assist births in weekday than weekend.  No significant difference in the days prior to weekend (Thursday and Fridays).







































